Offline Mode for Corporate Mobile Client Security Architecture

### 2016-06-16 docs/papers/2016JGC/2016JGC.tex
more...

% text
\usepackage[utf8x]{inputenc}

@@

% reference and bib issues
\usepackage[numbers]{natbib}
\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}
\hypersetup{breaklinks=true}
\urlstyle{same}
\Urlmuskip=0mu plus 1mu\relax

@@

\section{Introduction}
\label{sec_intro}
Cloud computing is a new rapidly evolving paradigm in the world of distributed networking and computation. The basic features of the cloud environment is providing the elastic, on-demand and secure services for the end-users. While the first two requirements are rather well conceptualized and supported by the majority of the cloud platforms in use, security is a serious concern of the cloud providers and governmental organizations as well as academia and research community \cite{csa2016,higashi2015,gartner2015}. Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized leading to obscure and incomplete security paradigms and solutions.

@@

Suppose that a SME uses CASB in order to protect data at rest, i.e. while stored on the server-side, in transit, communicating with server, and, in use, while the client is connected to the network. In this case, when the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools cannot help and the mobile app has to secure itself with its own limited resources. Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection. For example, offline mode does not allow performing the extensive computation and encryption on the mobile device.
Observing the above described landscape, this paper outlines the concept of the offline mobile client security. This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing \cite{galibus2007} and ABE encryption \cite{goyal2006attribute}. Moreover, our proposed approach proposal includes the usage of the user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to highlight possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack. The key expiration period is safely incorporated into the proposed system solution in order to enhance security. Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors. The main target of our proposed solution is to provide a maximum defense at the minimal resource cost.

@@

\begin{enumerate}
	\item \textbf{Master-key (MK)} which is kept safely on server and accessible only for the domain administrator and is given by $MK = (t_1, t_2,...,t_n, y)$, where the values $t_i$ are randomly selected from the huge group $Zp$. They are the private keys corresponding to the group attributes. Note, that this is different from the usual PK encryption: the private keys are controlled by the admin and not by the users.
	\item \textbf{Public key (PK)} depends on the master key values and is kept in the clear allowing users to access the information: $PK = (g^{t_1}, g^{t_2}",...,g^{t_n}, e(g,g)^y)$, Here $e(g,g)$ is the bilinear pairing function corresponding to an elliptic curve.
	\item Secret user \textbf{KEY\_SET} depends on his attribute set. Here each $D_i$ (\textbf{GROUP\_KEY}) serves for decryption of the data of a single group of users, for example, related to some project:

@@

Since the underlying sharing scheme is perfect \cite{galibus2008some}, the adversary cannot get any information of the KEY\_SET\_KEY unless he possessing all 4 key parts. Here the values PASS and PIN are predefined similar to the construction in \cite{galibus2015mobile}.

@@

\begin{equation}\label{eq:eq1}
  \left\{
  \begin{array}{@{}ll@{}}
    S = s \textrm{ mod } p_0
    \\
    S = s_1 \textrm{ mod } p_1
  \end{array}\right.
\end{equation}

According to the CRT:
\begin{equation}\label{eq:eq2}
S \equiv {s_{1}p_{0}{p_0}^{-1}}_{p1} + {s_{1}p_{1}{p_1}^{-1}}_{p0} + \textrm{ mod } p_0p_1
\end{equation}

@@

Model Order Selection is an effective signal processing technique for several applications, allowing separating the only noise components from the principal components applying a rank reduction of the data. Applying MOS to the analysis of user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. MOS for threat intelligence requires that the target features, which can be user operations, are modeled as a matrix composed by the number of occurrences by location and time. For example, one element can represent the number of file readings on folder $m$ during the minute $n$.

The MOS schemes can rely on correlation or covariance analysis, where the correlation analysis identifies anomalies on sparse or subtle number of file operations, while the MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period.

Classical approaches to model order selection require the computation of the sample covariance matrix Vxx and of its eigenvalues, obtained from the measurement matrix X of the zero mean samples given by

The set of obtained vectors zm composes the zero mean matrix Z, then the covariance matrix Vxx can be calculated as follows

Vxx =1NZZT

For MOS based on correlation analysis, instead of calculate the zero mean samples, to obtain the correlation matrix Rxx , it is required, for each variable, to make the standard deviation unitary as follows

The set of vectors sm composes the matrix S, then the correlation matrix Rxx can be calculated via 
. Once the Vxx or Rxx have been obtained for MOS in order to anomaly detection, for the sake of simplicity, we refer to Vxx and Rxx as a matrix C. Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to , in order to obtain the vector of eigenvalues e, as following:

The eigenvalues should be sorted in descending order, as defined by , to make possible the selection of the first eigenvalue in the obtained sequence, represented by, which is the largest eigenvalue of the data evaluated for attack detection.

The process of obtaining the and the matrix C(t), finding the largest eigenvalue for each t-th time frame, should be repeated until t = T, in order to obtain the largest eigenvalue of all time frames, as presented by

E =

Since   , then the first line of the matrix E contains the largest eigenvalues of each t-th time frame, which is the expected input for MOS schemes and can be expressed as 
emax=E{:,1}

Once obtained the largest eigenvalues of each q-th time frame, it is possible to apply a selected MOS scheme to estimate the model order $\hat{d}$, which is the estimated number of time frames with malicious behavior. Therefore, emax is used as input parameter for MOS schemes, according to the equation.

Note that some MOS schemes may also require the amount of time that compose a time frame, such as . For more information about MOS schemes, interested readers are referred to \cite{da2009comparison}.

@@

... add some text

@@

Steal the user keys one-by-one while the app performs decryption, if he is outsider;
Use expired key to decrypt the files, if he is insider;
Does not have possibility to perform offline dictionary attack and cryptanalysis, i.e. does not have the possibility to verify user credentials apart from entering them in the client app;
Steal the encrypted files but not the decrypted ones.

\subsubsection{Security analysis}
\label{sec_sec_analysis}
Offline mode
In the offline mode the client does not store the user password (see section 3 for details), i.e. no information about the password leaks, and therefore there is no possibility for the malicious user to check if the password he is trying to enter is correct or not. The only possible scenario for the information leakage in the case of the malicious outsider is if:
The hacker steals all the parts of KEY\_SET\_KEY;
The hacker steals the KEY\_SET;
The hacker try the brute force offline dictionary attack on all the previous values, they have to belong to ONE TIME SESSION (the values of ”a” and ”b” belong to one period of time i.e. TIME, DEV\_PASS, KEY\_SET);
Steal the permanent FILE\_KEY (this is protected by the KEY\_SET);
Steal the file and try to decrypt it with offline dictionary attacks.
        Still, the hacker has to get 4 values: TIME, DEV\_PASS, KEY\_SET, FILE\_KEY - from one session. At the same time he should try the offline dictionary attack on PIN+PASS. Moreover, the 4 values provide access only to 1 single file. So practically, it is very difficult to perform such attack due to the key expiration period. In accordance with the conditions 1) and 3) of our adversary model, the above actions are not possible for such adversary in the conditions presented.

The temporary nature of all parameters obliges the user to connect to server when necessary and prevents the malicious actions from the user side. The only possible scenario when malicious or expired user wishes to prolong his old credentials is:
He has to steal the DEV\_PASS and TIME synchronized with his credentials;
He has to be able to combine;
He has to steal the KEY\_SET synchronized with his credentials;
He has to steal the protected FILE\_KEY (also synchronized) and encrypted file;
He has to do everything without the client (because the client checks the TIME and renews the PUBLIC\_SHARE\_KEY).
Basically, for a malicious user there is practically no way to use the client with the old keys. The fact that the client does not contain any data to be checked or validated does not prevent the user from seeing the contents of decrypted files, but with the wrong password the decrypted files will be different from the original ones.

The client still has to count the tries (to prevent the hacker to perform brute force attack) within one session. In accordance with the conditions 2) and 4) of the adversary model, the above actions are not possible for such adversary.
Online mode
In the online mode the protection of the mobile system is backed up by the communication with server. The key expiry period and the usage of J-PAKE \cite{hao2010j,hao2008password} guarantees that the sensitive data sent over the network cannot leak. The detailed security analysis of J-PAKE is presented in \cite{toorani2014security}. The data sent in the clear i.e. encrypted files and the public keys of ABE does not need to be protected additionally as the AES and ABE guarantee its security.

@@

\subsubsection{An attacker uses a valid password to perform operations on a bulk of files}
\label{sec_attacker_valid}
The session time defines the period when operations can be performed until the next session renewing. During this period, it is still necessary to identify attacks and malicious behavior on file operations, in order to avoid fast attacks to perform unauthorized access to information or data modification. Some attacks present behavioral patterns based on abrupt number of operations, such as the ransomware attack, which is a growing attack \cite{McAfee2015} that blocks the access to valuable resources and requires a payment in order to unblock the content. The access to the resources can be blocked by the attacker through some techniques, when the content is encrypted by the attacker, the ransomware attack can be called cryptoransomware \cite{kaspersky2014}.
MOS schemes based on covariance analysis are effective to reveal abrupt changing of behaviors over time \cite{tenorio2013greatest}, making possible to identify intense malicious behaviors on offline mode of mobile clients, such in case of ransomware attack or bulk access to sensitive data.
The large number of operations over time is a well-known pattern of some attacks, due to the efforts on security measures to make the attacks infeasible over time. In this context, the operations can also be evaluated in contrast to the estimated required time for operations done by legitimate behaviors, such as the evaluation of the mean time between operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities.
Sparse or subtle file operations, with low number of operations distributed over different files or directories, during short period of time can indicate anomalies in contrast to the required time for legitimate directory navigation. MOS and correlation analysis can be suitable if applied to evaluate the time and location of operations, in order to identify unreachable navigation, if compared to legitimate navigation
The MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period. The eigenvalue analysis based on the correlation matrix highlights massive or concentrated operations over time or folder location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.
This threat scenario, where an attacker uses a valid password and session to perform operations on a bulk of files, can have its steps described as:
The hacker has access to the mobile client and is able to perform operations;
The session time is valid;
The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
The client incrementally append each operation attempt time into the logging;
The MOS module evaluates the logging of legitimate operations, applying correlation analysis to identify anomalies on sparse or subtle number of file operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities;
The MOS module evaluates the logging of legitimate operations, applying covariance analysis to identify abnormalities caused by massive operations during the session time;

\subsubsection{Usage of expired password to perform unauthorized operations }
\label{sec_attacker_expired}
In the offline mode, the session time is used to restrict the operations during a specified period, although it is possible to manipulate the current time in mobile clients, to emulate a period in which the session was valid. The log analysis by MOS can deal with this kind of threat, through the incremental logging of the time when each operation was performed, followed by the behavioral evaluation of operations over time. 
The incremental logging assumes that new logged operations shall have equal or bigger time than the last logged operation, the violation of this rule means that the system is out of sync and indicates a malicious behavior. Additionally, a large amount or sparse operation performed at the same time, or during a short period, can indicate the use of backtrack techniques to maintain a valid session during necessary time to perform an attack. Massive, subtle or sparse malicious operation performed during a valid session time can be identified by MOS schemes based on covariance and correlation analysis.
Applying MOS to the analysis of the time between user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. The MOS based on correlation analysis identifies anomalies on sparse or subtle number of file operations, since the eigenvalue analysis based on the correlation matrix highlights the unexpected number of sparse (such as file operations on diverse folders) or subtle operations. Consequently, the result of the eigenvalue analysis is applied to MOS schemes, in order to identify the occurrence of malicious behaviors during the valid session.
The MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period. The eigenvalue analysis based on correlation matrix highlights massive or concentrated operations over time or location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.
This threat scenario, where the attacker uses expired password to perform unauthorized operations, can have its steps described as:
The hacker steals the operating system;
The hacker modifies the time of the operating system to a period when the session was valid;
The hacker has access to the mobile client and is able to perform operations;
The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
The client incrementally append each operation attempt time into the logging;
The client verifies if one logged time is older than the last operation time. If it is true, the MOS module classifies the evaluated operation as malicious;
The MOS module evaluates the logging of legitimate operations, applying correlation analysis to identify anomalies on sparse or subtle number of file operations;
The MOS module evaluates the logging of legitimate operations, applying covariance analysis to identify abnormalities caused by massive operations during the session time;

\subsubsection{Data Modeling for Behavioral Analysis }
\label{sec_data}
 MOS schemes are used in order to identify anomalous behavior that can indicate an attack and be used to prevent or avoid attack propagation. Therefore, it is necessary to analyze the data that can be collected from user operations on mobile client, to identify features that can be modeled and submitted to MOS schemes, according to described in Section 5.4.
The selected features shall be modeled as matrices which represents a signal superposition containing noise, legitimate and malicious behavior \cite{tenorio2013greatest}, grouped into time frames , where M defines the decomposition of a selected feature, N defines the time decomposition and represents the number of occurrences of the feature m during the time n.

@@

file decryption, 
decrypted file reading 
decrypted file execution, 
Therefore, it is necessary to generate three matrices for malicious behaviors analysis, such as: 
massive file access, which can reveal data leakage and be identified by MOS schemes based on covariance analysis; 
low file access into several folders, characterized by sparse operations that can reveal unreachable navigation performed by automated file accesses in order to avoid the massive file access characterization. 
Malicious sparse file accesses can be identified by MOS schemes base correlation analysis.
File Update (Time and File System Location), i.e. writing operations into selected files in offline mode, writing the data stored on the mobile client. The update feature can be decomposed into:
 file encryption 
decrypted file writing, 
Therefore, it is necessary to generate two matrices for malicious behaviors analysis, such as: 
massive file update, which can reveal ransomware or similar attacks and be identified by MOS schemes based on covariance analysis; 
low number of file update into several folders, characterized by sparse operations that can reveal unreachable navigation performed by automated file accesses in order to avoid the massive file access characterization. Malicious sparse file accesses can be identified by MOS schemes base correlation analysis.
File Download (Start Time, End Time and File System Location), i.e download requests in online mode, evaluated by the mobile client. The file download feature shall be modeled as the matrix of number downloads by file location over time, in order to perform malicious behaviors analysis, such as:
massive data leakage or similar attacks, which can be identified by MOS schemes based on covariance analysis; 
low number of file download from several folders, characterized by sparse operations, which can reveal unreachable navigation performed by automated file download in order to avoid the massive file download characterization. Malicious sparse file download can be identified by MOS schemes base correlation analysis.

>>>>

% text
\usepackage[utf8x]{inputenc}
\usepackage{enumitem}
\newtheorem{thm}{Scenario}

% math
\usepackage{amsfonts}
\usepackage{amsmath}

@@

% reference and bib
\usepackage[numbers]{natbib}
\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}
\hypersetup{breaklinks=true}
\urlstyle{same}
\Urlmuskip=0mu plus 1mu\relax

@@

\section{Introduction}
\label{sec_intro}
Cloud computing is a new rapidly evolving paradigm in the world of distributed networking and computation. The basic features of the cloud environment is providing the elastic, on-demand and secure services for the end-users. While the first two requirements are rather well conceptualized and supported by the majority of the cloud platforms in use, security is a serious concern of the cloud providers and governmental organizations as well as academia and research community \cite{csa2016,higashi2015,gartner2015}. Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized leading to obscure and incomplete security paradigms and solutions.

@@

Suppose that a SME uses CASB in order to protect data at rest, i.e. while stored on the server-side, in transit, communicating with server, and, in use, while the client is connected to the network. In this case, when the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools cannot help and the mobile app has to secure itself with its own limited resources. Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection. For example, offline mode does not allow performing the extensive computation and encryption on the mobile device.
Observing the above described landscape, this paper outlines the concept of the offline mobile client security. This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing \cite{galibus2007} and ABE encryption \cite{goyal2006attribute}. Moreover, our proposed approach proposal includes the usage of the user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to highlight possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack. The key expiration period is safely incorporated into the proposed system solution in order to enhance security. Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors. The main target of our proposed solution is to provide a maximum defense at the minimal resource cost.

@@

\begin{enumerate}
	\item \textbf{Master-key (MK)} which is kept safely on server and accessible only for the domain administrator and is given by $MK = (t_1, t_2, \cdots ,t_n, y)$, where the values $t_i$ are randomly selected from the huge group $Zp$. They are the private keys corresponding to the group attributes. Note, that this is different from the usual PK encryption: the private keys are controlled by the admin and not by the users.
	\item \textbf{Public key (PK)} depends on the master key values and is kept in the clear allowing users to access the information: $PK = (g^{t_1}, g^{t_2}", \cdots ,g^{t_n}, e(g,g)^y)$, Here $e(g,g)$ is the bilinear pairing function corresponding to an elliptic curve.
	\item Secret user \textbf{KEY\_SET} depends on his attribute set. Here each $D_i$ (\textbf{GROUP\_KEY}) serves for decryption of the data of a single group of users, for example, related to some project:

@@

Since the underlying sharing scheme is perfect \cite{galibus2008some}, hence the adversary cannot get any information of the KEY\_SET\_KEY unless he possessing all 4 key parts. Here the values PASS and PIN are predefined similar to the construction in \cite{galibus2015mobile}.

@@

\begin{equation}\label{eq:eq01}
  \left\{
  \begin{array}{@{}ll@{}}
    S = s \textrm{ mod } p_0
    \\
    S = s_1 \textrm{ mod } p_1
  \end{array}\right.
\end{equation}

According to the CRT:
\begin{equation}\label{eq:eq02}
S \equiv {s_{1}p_{0}{p_0}^{-1}}_{p1} + {s_{1}p_{1}{p_1}^{-1}}_{p0} + \textrm{ mod } p_0p_1
\end{equation}

@@

Model Order Selection is an effective signal processing technique for several applications, allowing separating the only noise components from the principal components applying a rank reduction of the data. Applying MOS to the analysis of user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. MOS for threat intelligence requires that the target features, which can be user operations, are modeled as a matrix composed by the number of occurrences by location and time, and grouped into $Q$ time frames. Therefore, the framework considers the time variations of the matrix $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the occurrence of malicious behaviors. For example, one element of $\mathbf{X}^{(q)}$ can represent the number of file readings on folder $m$ during the minute $n$, from file operations logged by the client app.

The MOS schemes can rely on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, where the former is useful to identify abnormalities caused by large amounts of operations during a period, while the latter is applied to identify anomalies on sparse or subtle number of file operations.

Classical approaches to model order selection require the computation of the sample covariance matrix $\mathbf{\hat{R}}_{yy}^{(q)}$ and of its eigenvalues, obtained from the measurement matrix $\mathbf{X}$ of the zero mean samples given by

\begin{equation}\label{eq:eq03}
\mathbf{y}_{m}^{(q)} = \mathbf{x}_{m}^{(q)} - \bar{\mathbf{x}}_{m}^{(q)}.
\end{equation}

The set of obtained vectors $\mathbf{y}_{m}^{(q)}$ composes the zero mean matrix $\mathbf{Y}^{(q)}$, then the sample covariance matrix $\mathbf{\hat{R}}_{yy}^{(q)}$ can be calculated as follows

\begin{equation}\label{eq:eq04}
\mathbf{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\mathbf{Y}^{(q)}\mathbf{Y}^{(q)^{\rm T}}.
\end{equation}

For MOS based on sample covariance of zero mean and unitary standard deviation, it is required, for each variable, to make the standard deviation unitary as follows

\begin{equation}\label{eq:eq05}
\mathbf{z}_{m}^{(q)} = \frac{\mathbf{x}_{m}^{(q)} - \bar{\mathbf{x}}_{m}^{(q)}}{\mathbf{\sigma}_{m}^{(q)}}.
\end{equation}

The set of vectors $\mathbf{z}_{m}^{(q)}$ composes the matrix $\mathbf{Z}^{(q)}$, then the sample covariance matrix $\mathbf{\hat{R}}_{zz}^{(q)}$ can be calculated via

\begin{equation}\label{eq:eq06}
\mathbf{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\mathbf{Z}^{(q)}\mathbf{Z}^{(q)^{\rm T}}.
\end{equation}

Once the $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ have been obtained for MOS in order to anomaly detection, for the sake of simplicity, we refer to $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ as a matrix $\mathbf{C}$. Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to $\mathbf{C}^{(q)} = \mathbf{V}^{(q)}\mathbf{\Lambda}^{(q)}\mathbf{V}^{(q)^{\rm T}},$, in order to obtain the vector of eigenvalues $e$, as following:

\begin{equation}\label{eq:eq07}
\mathbf{e}^{(q)} = \rm diag(\mathbf{\Lambda}^{(q)}),
\end{equation}

The eigenvalues should be sorted in descending order, as defined by $\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > \cdots > \lambda_{m}^{(q)}$, to make possible the selection of the first eigenvalue in the obtained sequence, represented by $\lambda_{1}^{(q)}$, which is the largest eigenvalue of the data evaluated for attack detection.

The process of obtaining the $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$ and the matrix $\mathbf{C}^{(q)}$, finding the largest eigenvalue for each $q$-th time frame, should be repeated until $q = Q$, in order to obtain the largest eigenvalue of all time frames, as presented by

\begin{equation}\label{eq:eq08}
\mathbf{E} =
\begin{bmatrix}
  \lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
  \lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
  \lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
  \vdots & \vdots & \ddots & \vdots  \\
  \lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
\end{bmatrix}.
\end{equation}

Since $\lambda_1^{(q)} > \lambda_2^{(q)} > \lambda_3^{(q)} > \cdots > \lambda_{m-1}^{(q)} > \lambda_m^{(q)}$, then the first line of the matrix $\mathbf{E}$ contains the largest eigenvalues of each $q$-th time frame, which is the expected input for MOS schemes and can be expressed as

\begin{equation}\label{eq:eq09}
\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} \cdots \lambda_1^{(Q)}]
\end{equation}

Once obtained the largest eigenvalues of each $q$-th time frame, it is possible to apply a selected MOS scheme to estimate the model order $\hat{d}$, which is the estimated number of time frames with malicious behavior. Therefore, $\boldsymbol{e}_{\rm max}$ is used as input parameter for MOS schemes, according to the equation

\begin{equation}\label{eq:eq09}
\hat{d} = \textrm{MOS}(\boldsymbol{e}_{\rm max})
\end{equation}

Note that some MOS schemes may also require the amount of time that compose a time frame, such as $\hat{d} = \textrm{MOS}(\boldsymbol{e}_{\rm max}, M)$. For more information about MOS schemes, interested readers are referred to \cite{da2009comparison}.

@@

This section provides the detailed analysis of the results the proposed approach regarding security and complexity analysis.

@@

\begin{enumerate}
	\item Steal the user keys one-by-one while the app performs decryption, if he is outsider;
	\item Use expired key to decrypt the files, if he is insider;
	\item Does not have possibility to perform offline dictionary attack and cryptanalysis, i.e. does not have the possibility to verify user credentials apart from entering them in the client app;
	\item Steal the encrypted files but not the decrypted ones.
\end{enumerate}

\subsubsection{Security analysis}
\label{sec_sec_analysis}

\textbf{Offline Mode}\\

In the offline mode the client does not store the user password (see Section \ref{sec_offline_mode_workflow} for details), i.e. no information about the password leaks, and therefore there is no possibility for the malicious user to check if the password he is trying to enter is correct or not. The only possible scenario for the information leakage in the case of the malicious outsider is if:

\begin{enumerate}[label=(\alph*)]
	\item The hacker steals all the parts of KEY\_SET\_KEY;
	\item The hacker steals the KEY\_SET;
	\item The hacker try the brute force offline dictionary attack on all the previous values, they have to belong to ONE TIME SESSION (the values of ”a” and ”b” belong to one period of time i.e. TIME, DEV\_PASS, KEY\_SET);
	\item Steal the permanent FILE\_KEY (this is protected by the KEY\_SET);
	\item Steal the file and try to decrypt it with offline dictionary attacks.
\end{enumerate}

Still, the hacker has to get 4 values from one session: TIME, DEV\_PASS, KEY\_SET, FILE\_KEY. At the same time he should try the offline dictionary attack on PIN+PASS. Moreover, the 4 values provide access only to 1 single file. So practically, it is very difficult to perform such attack due to the key expiration period. In accordance with the conditions 1 and 3 of our adversary model, the above actions are not possible for such adversary in the conditions presented.

The temporary nature of all parameters obliges the user to connect to server when necessary and prevents the malicious actions from the user side. The only possible scenario when malicious or expired user wishes to prolong his old credentials is:

\begin{enumerate}[label=(\alph*)]
	\item Steal the file and try to decrypt it with offline dictionary attacks.
	\item He has to be able to combine;
	\item He has to steal the KEY\_SET synchronized with his credentials;
	\item He has to steal the protected FILE\_KEY (also synchronized) and encrypted file;
	\item He has to do everything without the client (because the client checks the TIME and renews the PUBLIC\_SHARE\_KEY).
\end{enumerate}

Basically, for a malicious user there is practically no way to use the client with the old keys. The fact that the client does not contain any data to be checked or validated does not prevent the user from seeing the contents of decrypted files, but with the wrong password the decrypted files will be different from the original ones.

The client still has to count the tries (to prevent the hacker to perform brute force attack) within one session. In accordance with the conditions 2 and 4 of the adversary model, the above actions are not possible for such adversary.
\\
\\
\textbf{Online mode}\\

In the online mode the protection of the mobile system is backed up by the communication with server. The key expiry period and the usage of J-PAKE \cite{hao2010j,hao2008password} guarantees that the sensitive data sent over the network cannot leak. The detailed security analysis of J-PAKE is presented in \cite{toorani2014security}. The data sent in the clear i.e. encrypted files and the public keys of ABE does not need to be protected additionally as the AES and ABE guarantee its security.

@@

\begin{thm}
An attacker uses a valid password to perform operations on a bulk of files.
\end{thm}

The session time defines the period when operations can be performed until the next session renewing. During this period, it is still necessary to identify attacks and malicious behavior on file operations, in order to avoid fast attacks to perform unauthorized access to information or data modification. Some attacks present behavioral patterns based on abrupt number of operations, such as the ransomware attack, which is a growing attack \cite{McAfee2015} that blocks the access to valuable resources and requires a payment in order to unblock the content. The access to the resources can be blocked by the attacker through some techniques, when the content is encrypted by the attacker, the ransomware attack can be called cryptoransomware \cite{kaspersky2014}.

MOS schemes based on covariance analysis are effective to reveal abrupt changing of behaviors over time \cite{tenorio2013greatest}, making possible to identify intense malicious behaviors on offline mode of mobile clients, such in case of ransomware attack or bulk access to sensitive data.

The large number of operations over time is a well-known pattern of some attacks, due to the efforts on security measures to make the attacks infeasible over time. In this context, the operations can also be evaluated in contrast to the estimated required time for operations done by legitimate behaviors, such as the evaluation of the mean time between operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities.

Sparse or subtle file operations, with low number of operations distributed over different files or directories, during short period of time can indicate anomalies in contrast to the required time for legitimate directory navigation. MOS and correlation analysis can be suitable if applied to evaluate the time and location of operations, in order to identify unreachable navigation, if compared to legitimate navigation

The MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period. The eigenvalue analysis based on the correlation matrix highlights massive or concentrated operations over time or folder location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.

This threat scenario, where an attacker uses a valid password and session to perform operations on a bulk of files, can have its steps described as:

\begin{enumerate}[label=(\alph*)]
	\item The hacker has access to the mobile client and is able to perform operations;
	\item The session time is valid;
	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
	\item The client incrementally append each operation attempt time into the logging;
	\item The MOS module evaluates the logging of legitimate operations, applying correlation analysis to identify anomalies on sparse or subtle number of file operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities;
	\item The MOS module evaluates the logging of legitimate operations, applying covariance analysis to identify abnormalities caused by massive operations during the session time.
\end{enumerate}

\begin{thm}
Usage of expired password to perform unauthorized operations.
\end{thm}

In the offline mode, the session time is used to restrict the operations during a specified period, although it is possible to manipulate the current time in mobile clients, to emulate a period in which the session was valid. The log analysis by MOS can deal with this kind of threat, through the incremental logging of the time when each operation was performed, followed by the behavioral evaluation of operations over time.

The incremental logging assumes that new logged operations shall have equal or bigger time than the last logged operation, the violation of this rule means that the system is out of sync and indicates a malicious behavior. Additionally, a large amount or sparse operation performed at the same time, or during a short period, can indicate the use of backtrack techniques to maintain a valid session during necessary time to perform an attack. Massive, subtle or sparse malicious operation performed during a valid session time can be identified by MOS schemes based on covariance and correlation analysis.

Applying MOS to the analysis of the time between user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. The MOS based on correlation analysis identifies anomalies on sparse or subtle number of file operations, since the eigenvalue analysis based on the correlation matrix highlights the unexpected number of sparse (such as file operations on diverse folders) or subtle operations. Consequently, the result of the eigenvalue analysis is applied to MOS schemes, in order to identify the occurrence of malicious behaviors during the valid session.

The MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period. The eigenvalue analysis based on correlation matrix highlights massive or concentrated operations over time or location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.

This threat scenario, where the attacker uses expired password to perform unauthorized operations, can have its steps described as:

\begin{enumerate}[label=(\alph*)]
	\item The hacker steals the operating system;
	\item The hacker modifies the time of the operating system to a period when the session was valid;
	\item The hacker has access to the mobile client and is able to perform operations;
	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
	\item The client incrementally append each operation attempt time into the logging;
	\item The client verifies if one logged time is older than the last operation time. If it is true, the MOS module classifies the evaluated operation as malicious;
	\item The MOS module evaluates the logging of legitimate operations, applying correlation analysis to identify anomalies on sparse or subtle number of file operations;
	\item The MOS module evaluates the logging of legitimate operations, applying covariance analysis to identify abnormalities caused by massive operations during the session time;
\end{enumerate}

\subsubsection{Data Modeling for Behavioral Analysis }
\label{sec_data}

MOS schemes are used in order to identify anomalous behavior that can indicate an attack and be used to prevent or avoid attack propagation. Therefore, it is necessary to analyze the data that can be collected from user operations on mobile client, to identify features that can be modeled and submitted to MOS schemes, according to described in Section \ref{sec_mos}.

The selected features shall be modeled as matrices which represents a signal superposition containing noise, legitimate and malicious behavior \cite{tenorio2013greatest}, grouped into time frames $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, 2, 3, \ldots, Q$ , where $M$ defines the decomposition of a selected feature, $N$ defines the time decomposition and represents the number of occurrences of the feature $m$ during the time $n$.

@@

file decryption, 
decrypted file reading 
decrypted file execution,
